---
title: "brgedata -- data R package with three omic data-set and exposome data-set from the same Spanish pupulation"
author:
- name: Carles Hernandez-Ferrer
  affiliation: 
  - &isg Barcelona Institute for Global Health
- name: Carlos Ruiz-Arenas
  affiliation: *isg
- name: Juan R. Gonzalez
  affiliation: *isg
  email: juanr.gonzalez@isglobal.org
date: "`r doc_date()`"
package: "`r pkg_ver('brgedata')`"
output:
  BiocStyle::html_document
abstract: |
  This document indicates the different data-sets included in `brgedata`. The description includes the technology used to create the and the number of samples and features in each one.
vignette: >
  %\VignetteIndexEntry{brgedata -- data R package with three omic data-set and exposome data-set from the same Spanish pupulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Getting started

`brgedata` includes a collection of BRGE omic and exposome data from the same cohort. The diferent objects guarantees a minimum of samples in common between all sets.

Data available in this R package:

| Data Type     | Number of Samples | Number of Features | Technology                      | Object Name  | Class             |
| :------------ | ----------------: | -----------------: | :------------------------------ | :----------- | :---------------- |
| Exposome      | 110               | 15                 |                                 | `brge_expo`  | [`ExposomeSet`](https://github.com/isglobal-brge/rexposome) |
| Transcriptome | 75                | 67528              | Affymetrix HTA 2.0              | `brge_gexp`  | [`ExpressionSet`](https://bioconductor.org/packages/release/bioc/html/Biobase.html) |
| Methylome     | 20                | 476946             | Illumina Human Methylation 450K | `brge_methy` | [`GenomicRatioSet`](https://bioconductor.org/packages/release/bioc/html/minfi.html) |
| Proteome      | 90                | 47                 |                                 | `brge_prot`  | [`ExpressionSet`](https://bioconductor.org/packages/release/bioc/html/Biobase.html) |

`sex` and `age` was included as phenotipic data in each set. Moreover, the `ExposomeSet` includes *asthma status* and *rhinitis status* of each sample.

# Data Resources

## Exposome Data

To load the **exposome data**, stored in an `ExposomeSet`, run the follow commands:

```{r load_exposome, message=FALSE}
data("brge_expo", package = "brgedata")
brge_expo
```

The summary of the data contained by `brge_expo`:

| Data Type     | Number of Samples | Number of Features | Technology                      | Object Name  | Class             |
| :------------ | ----------------: | -----------------: | :------------------------------ | :----------- | :---------------- |
| Exposome      | 110               | 15                 |                                 | `brge_expo`  | [`ExposomeSet`](https://github.com/isglobal-brge/rexposome) |


## Transcriptome Data

To load the **transcriptome data**, saved in an `ExpressionSet`, run the follow commands:

```{r load_transcriptome, message=FALSE}
data("brge_gexp", package = "brgedata")
brge_gexp
```

The summary of the data contained by `brge_gexp`:

| Data Type     | Number of Samples | Number of Features | Technology                      | Object Name  | Class             |
| :------------ | ----------------: | -----------------: | :------------------------------ | :----------- | :---------------- |
| Transcriptome | 75                | 67528              | Affymetrix HTA 2.0              | `brge_gexp`  | [`ExpressionSet`](https://bioconductor.org/packages/release/bioc/html/Biobase.html) |


## Methylome Data

To load the **methylation data**, encapsulated in a `GenomicRatioSet`, run the follow commands:

```{r load_methylome, message=FALSE}
data("brge_methy", package = "brgedata")
brge_methy
```

The summary of the data contained by `brge_methy`:

| Data Type     | Number of Samples | Number of Features | Technology                      | Object Name  | Class             |
| :------------ | ----------------: | -----------------: | :------------------------------ | :----------- | :---------------- |
| Methylome     | 115               | 476946             | Illumina Human Methylation 450K | `brge_methy` | [`GenomicRatioSet`](https://bioconductor.org/packages/release/bioc/html/minfi.html) |


## Proteome Data

To load the **protein data**, stored in an `ExpressionSet`, run the follow commands:

```{r load_proteome, message=FALSE}
data("brge_prot", package = "brgedata")
brge_prot
```

The summary of the data contained by `brge_prot`:

| Data Type     | Number of Samples | Number of Features | Technology                      | Object Name  | Class             |
| :------------ | ----------------: | -----------------: | :------------------------------ | :----------- | :---------------- |
| Proteome      | 90                | 47                 |                                 | `brge_prot`  | [`ExpressionSet`](https://bioconductor.org/packages/release/bioc/html/Biobase.html) |

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
